\chapter{Introduction}\label{introduction}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1.1 What We're Building}\label{what-were-building}

Welcome to this comprehensive guide on building a modern educational
platform from scratch. Throughout this book, we'll create
\textbf{Afidna} --- a fully-featured learning management system (LMS)
designed for Islamic scholarly content.

By the end of this journey, you'll have built:

\begin{itemize}
\tightlist
\item
  \textbf{A content delivery system} with hierarchical organization
  (Tracks → Series → Lessons → Videos)
\item
  \textbf{User authentication} with secure session management
\item
  \textbf{Progress tracking} that saves watch time and quiz scores
\item
  \textbf{An interactive quiz system} with instant feedback
\item
  \textbf{A video player} integrated with YouTube's API
\item
  \textbf{An admin panel} for content management
\item
  \textbf{A global search} across all content
\item
  \textbf{A responsive UI} supporting both LTR and RTL languages
\end{itemize}

This isn't a toy project or a simple tutorial app. We're building
production-ready software that can serve thousands of users.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1.2 Why This Tech Stack?}\label{why-this-tech-stack}

\subsection{The Traditional Approach (and Its
Problems)}\label{the-traditional-approach-and-its-problems}

Many educational platforms are built with: - \textbf{Python/Django} or
\textbf{Node/Express} for the backend - \textbf{React} or \textbf{Vue}
for the frontend - \textbf{PostgreSQL} or \textbf{MySQL} for the
database - \textbf{Redis} for sessions - \textbf{Nginx} for reverse
proxy

This stack works, but it comes with complexity: - Multiple services to
manage - High memory footprint - Complex deployment - Expensive hosting

\subsection{Our Modern Approach}\label{our-modern-approach}

We're taking a different path --- the \textbf{``Compact Monolith''}
philosophy:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3929}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4286}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.1786}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Component
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Our Choice
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Why
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Runtime} & Bun & 4x faster than Node.js, built-in bundler,
native TypeScript \\
\textbf{Framework} & SvelteKit & SSR + SPA in one, excellent DX, small
bundle size \\
\textbf{Database} & SQLite & Zero configuration, single file,
surprisingly capable \\
\textbf{ORM} & Drizzle & Type-safe, lightweight, great SQLite support \\
\textbf{UI} & DaisyUI + Tailwind & Beautiful components, rapid
development \\
\textbf{Auth} & Session-based & Simple, secure, no JWT complexity \\
\end{longtable}

\subsection{The Result}\label{the-result}

\begin{itemize}
\tightlist
\item
  \textbf{Single process} to deploy
\item
  \textbf{\textless{} 100MB RAM} in production
\item
  \textbf{Sub-millisecond} database queries
\item
  \textbf{\$5/month} VPS is more than enough
\item
  \textbf{Full-stack TypeScript} --- same language everywhere
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1.3 Prerequisites}\label{prerequisites}

Before diving in, make sure you have:

\subsection{Technical Knowledge}\label{technical-knowledge}

\begin{itemize}
\tightlist
\item
  \textbf{JavaScript/TypeScript}: Comfortable with modern ES6+ syntax
\item
  \textbf{HTML/CSS}: Basic understanding of web markup
\item
  \textbf{Command Line}: Comfortable with terminal basics
\item
  \textbf{Git}: Basic version control knowledge
\end{itemize}

\subsection{Not Required (We'll Cover
These)}\label{not-required-well-cover-these}

\begin{itemize}
\tightlist
\item
  Svelte/SvelteKit experience
\item
  SQLite knowledge
\item
  Backend development experience
\item
  DevOps skills
\end{itemize}

\subsection{Software Requirements}\label{software-requirements}

\begin{itemize}
\tightlist
\item
  \textbf{Operating System}: Linux, macOS, or Windows (WSL recommended)
\item
  \textbf{Node.js}: Version 18+ (for some tooling)
\item
  \textbf{Bun}: We'll install this together
\item
  \textbf{Code Editor}: VS Code or Cursor recommended
\item
  \textbf{Git}: For version control
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1.4 Project Overview}\label{project-overview}

\subsection{High-Level Architecture}\label{high-level-architecture}

\begin{lstlisting}
┌─────────────────────────────────────────────────────────┐
│                      Client (Browser)                    │
│  ┌─────────────┐  ┌─────────────┐  ┌─────────────┐     │
│  │  Homepage   │  │   Lessons   │  │    Admin    │     │
│  └─────────────┘  └─────────────┘  └─────────────┘     │
└─────────────────────────┬───────────────────────────────┘
                          │ HTTP/HTTPS
┌─────────────────────────┴───────────────────────────────┐
│                    SvelteKit Server                      │
│  ┌─────────────┐  ┌─────────────┐  ┌─────────────┐     │
│  │    Pages    │  │     API     │  │    Auth     │     │
│  │   (SSR)     │  │  Endpoints  │  │  Middleware │     │
│  └─────────────┘  └─────────────┘  └─────────────┘     │
│                          │                               │
│  ┌───────────────────────┴────────────────────────┐    │
│  │                 Drizzle ORM                     │    │
│  └───────────────────────┬────────────────────────┘    │
└──────────────────────────┼──────────────────────────────┘
                           │
      ┌────────────────────┴────────────────────┐
      │                                          │
┌─────┴─────┐                            ┌──────┴──────┐
│ users.db  │                            │ content.db  │
│ (R/W)     │                            │ (Read-Only) │
│           │                            │             │
│ • users   │                            │ • tracks    │
│ • sessions│                            │ • series    │
│ • progress│                            │ • lessons   │
└───────────┘                            │ • videos    │
                                         └─────────────┘
\end{lstlisting}

\subsection{Key Design Decisions}\label{key-design-decisions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Dual Database Strategy}: Separate user data from content
\item
  \textbf{Hierarchical Content}: Tracks → Series → Lessons → Videos
\item
  \textbf{Human-First Editing}: Protect manual edits from automated
  imports
\item
  \textbf{Progressive Enhancement}: Works without JavaScript, better
  with it
\item
  \textbf{Mobile-First Design}: Responsive from the start
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1.5 Book Structure}\label{book-structure}

This book is organized into \textbf{10 parts}:

\subsection{Part I: Foundation \& Planning (Chapters
1-3)}\label{part-i-foundation-planning-chapters-1-3}

Setting up your environment and understanding the architecture.

\subsection{Part II: Backend Fundamentals (Chapters
4-7)}\label{part-ii-backend-fundamentals-chapters-4-7}

Database design, ORM setup, authentication, and API development.

\subsection{Part III: Frontend Development (Chapters
8-11)}\label{part-iii-frontend-development-chapters-8-11}

SvelteKit basics, UI components, and building core pages.

\subsection{Part IV: Advanced Features (Chapters
12-15)}\label{part-iv-advanced-features-chapters-12-15}

Video player, quiz system, progress tracking, and search.

\subsection{Part V: Admin Panel (Chapters
16-18)}\label{part-v-admin-panel-chapters-16-18}

Building a complete content management interface.

\subsection{Part VI: Content Pipeline (Chapters
19-20)}\label{part-vi-content-pipeline-chapters-19-20}

Automated content import and media management.

\subsection{Part VII: Testing \& Quality (Chapters
21-22)}\label{part-vii-testing-quality-chapters-21-22}

Testing strategies and code quality tools.

\subsection{Part VIII: Deployment (Chapters
23-26)}\label{part-viii-deployment-chapters-23-26}

Production builds, server setup, and monitoring.

\subsection{Part IX: Scaling \& Optimization (Chapters
27-28)}\label{part-ix-scaling-optimization-chapters-27-28}

Performance tuning and growth strategies.

\subsection{Part X: Appendices}\label{part-x-appendices}

Reference materials and troubleshooting guides.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{How to Use This Book}\label{how-to-use-this-book}

\subsection{If You're Following Along}\label{if-youre-following-along}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Type the code yourself} --- don't just copy-paste
\item
  \textbf{Run the code after each section} --- see it work
\item
  \textbf{Experiment} --- break things, fix them
\item
  \textbf{Take notes} --- document your learnings
\end{enumerate}

\subsection{If You're Using This as
Reference}\label{if-youre-using-this-as-reference}

\begin{itemize}
\tightlist
\item
  Each chapter is self-contained
\item
  Code examples are complete and runnable
\item
  The appendices have quick reference guides
\end{itemize}

\subsection{Code Repository}\label{code-repository}

All code from this book is available at:

\begin{lstlisting}
https://github.com/salahAit/afidna-Sveltekit
\end{lstlisting}

You can clone it to compare with your work or use it as a starting
point.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Let's Begin}\label{lets-begin}

In the next chapter, we'll dive deep into architecture design ---
understanding why we make certain decisions and how they impact our
application's future.

Get your development environment ready. We're about to build something
amazing.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Next Chapter}: \href{./chapter-02-architecture.md}{Chapter 2:
Architecture Design}
\end{quote}
